A star formation event in the leading arm of the Magellanic stream triggered by passage through the outer Galactic disk

### 2018-10-31 ms.tex
Update on Overleaf.
whitespace
todos
abstract and intro

% TODO
% - Sand, Seth, et al. Virgo: blue stars with nothing else
% - Mixture model in proper-motion space, Gaia only

% Proposal
% - IMACS?
% - Blue stars - bandpasses have bad Teff resolution
% - Better Teff, logg -> better age
% - Big FOV, blue throughput, southern, >4m

@@

\author[0000-0002-1793-3689]{David L. Nidever}
\affiliation{Department of Physics, Montana State University, P.O. Box 173840, Bozeman, MT 59717, USA}
\affiliation{National Optical Astronomy Observatory, 950 North Cherry Ave, Tucson, AZ 85719, USA}

\author[0000-0003-1680-1884]{Yumi Choi}
\affiliation{Department of Physics, Montana State University, P.O. Box 173840, Bozeman, MT 59717, USA}
\affiliation{Steward Observatory, University of Arizona, 933 North Cherry Avenue, Tucson AZ, 85721, USA}

\author[0000-0002-3569-7421]{Edward F. Schlafly}
\affiliation{Lawrence Berkeley National Laboratory, One Cyclotron Road, Berkeley, CA 94720, USA}

\author[0000-0003-2808-275X]{Douglas P. Finkbeiner}
\affiliation{Institute for Theory and Computation, Harvard-Smithsonian Center for Astrophysics, 60 Garden Street, MS-51, Cambridge, MA 02138, USA}

@@

We report the discovery of a young, $\lesssim 100~\textrm{Myr}$, metal-poor, $[\textrm{Fe}/\textrm{H}] \sim -1$, comoving stellar association in the halo of the Milky Way, $(R, z) \sim (24, 16)~\textrm{kpc}$.
The association --- \clustername\ --- is likely associated with the leading arm of the gas stream emanating from the Magellanic cloud system, but is located $\approx 60^\circ$ from the Large Magellanic Cloud (LMC) center, on the other side of the Milky Way disk relative to the LMC.
By assuming that the cluster is co-located with \hi gas in the stream, we directly measure the distance to the Magellanic stream, $d \approx 25~\textrm{kpc}$.
At this location relative to the LMC, the measured distance is inconsistent with predictions from models of the LMC/SMC interaction and infall into the Milky Way \todo{that don't account for ram pressure and gas interaction with MW disk}
The estimated age of \clustername\ is consistent with the time of last passage through the Galactic midplane.
We therefore conclude that this star-formation event was triggered by the last disk passage, which occurred at a Galactocentric radius $R \approx XX~\kpc$.
\todo{Further follow-up needed...spectroscopy}
\end{abstract}

@@

The origin and formation of the LA features in the MS is still uncertain.
Initial studies of the LA argued that the features closest to the \lmcsmc\ can be traced back to the SMC, thus implying that the gas in the LA was stripped from the SMC \citep{Putman:1998}. 
In later work, it has instead been argued that outer features of the LMC appear to lead directly into the LA-1 feature \citep{Nidever:2008}, implying an LMC origin for the leading arm.
Recent chemical abundance measurements along several sightlines passing through the LA again suggested the SMC origin \citep{Fox:2018,Richter:2018}. 
Recent hydrodynamical simulations of the MCs on the first infall to the Milky Way, however, showed that the repeated encounters between the MCs strip gas not only from the SMC, but also from the LMC, and this tidally stripped gas creates filamentary gas structures both in the LA and the Magellanic Stream \citep{Pardy:2018}.   
Whatever the origin of the gas, it is clear that tidal stripping by the Milky Way is required to form the leading arm \citep{Nidever:2008, Besla:2012}.
However, the LA features deviate from the predicted orbit of the \lmcsmc, implying that ram pressure or interactions with gas in the outer Milky Way disk may have removed orbital energy from the LA gas \citep[e.g.,][]{Bekki:2008}.

@@

In this \articlename, we report the discovery of a clustered association of young stars at the far edge of the LA-2 feature ($L_{\textrm{MS}} \sim 65^\circ$) that is unambiguously located far into the Galactic halo ($d_\odot \sim 30~\textrm{kpc}$) and therefore likely formed from gas in the leading arm of the Magellanic stream as it crossed the Galactic disk.
This provides a clear distance measurement to the leading arm of the MS to be compared with models of \lmcsmc\ interaction and infall into the Milky Way.

@@

% Notebook: Proper-motion-membership
\begin{figure}
\centering
\includegraphics[width=\textwidth]{figures/pm-model.pdf}
\caption{\todo{APW}
}
\label{fig:pm-members}
\end{figure}

>>>>

% TODO
% - Sand, Seth, et al. Virgo: blue stars with nothing else
% - Thank DFM
% - Thank Dylan and Nick @ SCC
% - Thank Flatiron

@@

\author[0000-0002-1793-3689]{David~L.~Nidever}
\affiliation{Department of Physics, Montana State University, P.O. Box 173840, Bozeman, MT 59717, USA}
\affiliation{National Optical Astronomy Observatory, 950 North Cherry Ave, Tucson, AZ 85719, USA}

\author[0000-0003-1680-1884]{Yumi~Choi}
\affiliation{Department of Physics, Montana State University, P.O. Box 173840, Bozeman, MT 59717, USA}
\affiliation{Steward Observatory, University of Arizona, 933 North Cherry Avenue, Tucson AZ, 85721, USA}

\author[0000-0002-8537-5711]{Timothy~Morton} % TODO: check orcID
\affiliation{Department of Astrophysical Sciences,
             Princeton University, Princeton, NJ 08544, USA}

\author[0000-0002-3569-7421]{Edward~F.~Schlafly}
\affiliation{Lawrence Berkeley National Laboratory, One Cyclotron Road, Berkeley, CA 94720, USA}

\author[0000-0003-2808-275X]{Douglas~P.~Finkbeiner}
\affiliation{Institute for Theory and Computation, Harvard-Smithsonian Center for Astrophysics, 60 Garden Street, MS-51, Cambridge, MA 02138, USA}

@@

We report the discovery of a young, $\sim 150~\textrm{Myr}$, metal-poor, $[\textrm{Fe}/\textrm{H}] \sim -1.2$, stellar association in the halo of the Milky Way, $(R, z) \sim (24, 16)~\textrm{kpc}$. % TODO: check with new dist.
The association --- \clustername\ --- is spatially near the leading arm of the gas stream emanating from the Magellanic cloud system, but is located $\approx 60^\circ$ from the Large Magellanic Cloud (LMC) center, on the other side of the Milky Way disk relative to the LMC.
By assuming that the cluster is co-located with \hi gas in the stream, we directly measure the distance to the Magellanic stream, $d \approx 30~\textrm{kpc}$.
At this location relative to the LMC, the measured distance is inconsistent with predictions from models of the LMC/SMC interaction and infall into the Milky Way that do not account for ram pressure and gas interaction with MW disk.
The estimated age of \clustername\ is consistent with the time of last passage through the Galactic midplane.
We therefore speculate that this star-formation event was triggered by the last disk midplane passage of the Magellanic stream gas, which occurred at a Galactocentric radius $R \approx XX~\kpc$.
However, most details of this idea remain a mystery: the Magellanic Stream is low column density, the Milky Way disk at this large radius has low surface density, and the relative velocity of the gas is large.
Spectroscopic follow-up is needed to confirm its kinematic association with the Magellanic Stream \hi, but either way \todo{this is an interesting thing.}
\end{abstract}

@@

The origin and formation of the LA features in the MS is still uncertain.
Initial studies of the LA argued that the features closest to the \lmcsmc\ can be traced back to the SMC, thus implying that the gas in the LA was stripped from the SMC \citep{Putman:1998}.
In later work, it has instead been argued that outer features of the LMC appear to lead directly into the LA-1 feature \citep{Nidever:2008}, implying an LMC origin for the leading arm.
Recent chemical abundance measurements along several sightlines passing through the LA again suggested the SMC origin \citep{Fox:2018, Richter:2018}.
Recent hydrodynamical simulations of the MCs on the first infall to the Milky Way, however, showed that the repeated encounters between the MCs strip gas not only from the SMC, but also from the LMC, and this tidally stripped gas creates filamentary gas structures both in the LA and the Magellanic Stream \citep{Pardy:2018}.
Whatever the origin of the gas, it is clear that tidal stripping by the Milky Way is required to form the leading arm \citep{Nidever:2008, Besla:2012}.
However, the LA features deviate from the predicted orbit of the \lmcsmc, implying that ram pressure or interactions with gas in the outer Milky Way disk may have removed orbital energy from the LA gas \citep[e.g.,][]{Bekki:2008}.

@@

In this \articlename, we report the discovery of a clustered association of young stars at the far edge of the LA-2 feature ($L_{\textrm{MS}} \sim 65^\circ$) that is unambiguously located far into the Galactic halo ($d_\odot \sim 30~\textrm{kpc}$) and therefore likely formed from gas in the leading arm of the Magellanic stream as it crossed the Galactic disk.
This provides a clear distance measurement to the leading arm of the MS to be compared with models of \lmcsmc\ interaction and infall into the Milky Way, and also provides an opportunity to study recent star formation in a regime unlike that of the Milky Way. \todo{or any other?}

@@

% Notebook: Proper-motion-membership
\begin{figure}
\centering
\includegraphics[width=\textwidth]{figures/pm-model.pdf}
\caption{\todo{APW}
Add proper motion vector?}
\label{fig:pm-members}
\end{figure}
